---
title: "Seasonal Surge and Off-Season Slump: Analyzing the Variance in Demand for the Toronto Island Ferry Over 2025-2026"
subtitle: "Month by month comparison of the ticket sales and redemptions"
author: 
  - Kyle Dong
thanks: "Code and data are available at: [https://github.com/KyleDong03/Toronto_Ferry_Sales](https://github.com/KyleDong03/Toronto_Ferry_Sales)."
date: today
date-format: long
abstract: "By analyzing the ticket sales and redemptions for the Toronto Island ferry Provided by the city of Toronto, this study aims to indentify how demand for the ferry changes throughout the year in spite of outside factors. The analysis finds that the ferry is more popular during the summer months, with June of 2026 outperforming June of 2025. This allows for the ferry services to optimize their staffing and vehicle scheduling over the course of a year, and through year to year comparisons it can be determined how effective new events are on increasing the Island's popularity."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
```


# Introduction

The Toronto Islands are the premier way to view the city's skyline. Thus it becomes a hub for travellers to visit, whether they are just visiting Toronto or are from the surrounding area. Understanding when these visits happen, and what can influence the demand allows for the City of Toronto, which operates the ferries, to best optimize their spending versus returns.

The City of Toronto keeps records of how many ferry tickets are sold, and redeemed, in real time with servers being updated every 15 minutes. By taking these data points and aggregating them by what month and year the sales occured, it was seen that the ferry is popular during the months of July and August, with November, December, January, February, and March being the least popular. Also when camparing between June 2025 and 2026, it can be seen that there is a significant difference in ticket sales and redemptions, which may be due to the multitude of new events on the Toronto Islands. 

This trend of having busy summers and off-peak winters can be crucial to the city as it allows them to have their infrastructure equiped for their peak seasons by taking advantage of the off-peak to reduce the operating budget and schedule maintenance while minimizing lost revenue. Understanding the correlation of new events and traffic to the Toronto Islands allows the City of Toronto to better decide what events to hold in the future and how they can optimize profits in such situations, allowing them to reinvest more into their infrastructure and personnel. 

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data provides a description of the data used in the analysis, @sec-results gives a detailed account of when the tickets are sold and redeemed, and @sec-sec-disc summarizes and discusses the results.






# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement
	
Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.



Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion {#sec-disc}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


